\LARGE \bf\vspace{-10mm} Face Recognition: Too Bias, or Not Too Bias?

### 2020-02-06 main.tex
version complete
updated supplemental citations

\usepackage{FG2019}
\usepackage{multirow}
\newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}
\usepackage[acronym]{glossaries}
\usepackage{acronym}
\usepackage{bigstrut}
% symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb,amsthm}
\usepackage{amsmath}% assumes amsmath package installed
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% \usepackage{hyperref}       % hyperlinks
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

\usepackage{url}            % simple URL typesetting
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{tikz}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
\usepackage{arydshln}
\usepackage{todonotes}
\usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}

@@

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[]{algorithm2e}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont

@@

\newacronym{sdm}{SDM}{signal detection model}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{nmse}{NMSE}{Normalized Mean Square Error}
\newacronym{det}{DET}{Detection Error Trade-off}
\newacronym{tp}{TP}{true-positive}
\newacronym{tpir}{TPIR}{true-positive identification rate}
\newacronym{frir}{FRIR}{false-reject identification rate}
\newacronym{fpir}{FRIR}{false-positive identification rate}

@@

\newacronym{tar}{TAR}{True Acceptance Rate}
\newacronym{far}{FAR}{False Acceptance Rate}
\newacronym{eer}{EER}{Equal Error Rate}
\newacronym{det}{DET}{Detection Error Trade-off}

@@

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}
\renewcommand{\arraystretch}{1.4}
\newcommand{\xmark}{\ding{53}}%

@@

\def\FGPaperID{41} % *** Enter the FG 2019 Paper ID here

@@

%use this in case of a single affiliation
%\author{\parbox{16cm}{\centering
%    {\large Huibert Kwakernaak}\\
%    {\normalsize
%    Faculty of Electrical Engineering, Mathematics and Computer Science, University of Twente, Enschede, The Netherlands\\}}
%    \thanks{This work was not supported by any organization.}% <-this % stops a space
%}

%use this in case of several affiliations
\author{\parbox{16cm}{\centering
    {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
    {\normalsize
    \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
    % $^1$ Northeastern University, Boston, MA, USA\\
    % $^2$ ISM Connect, Cambridge, MA, USA}}
    \thanks{}% <-this % stops a space
}

% \author{Joseph P Robinson$^1$, Yuncheng Li$^{2}$, Ning Zhang$^{2}$, Yun Fu$^{1}$, and Sergey Tulyakov$^{2}$\\
% \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$Snap Inc. \\\vspace{2mm}
% } %\\

\begin{document}
\begin{bibunit}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. Classic signal detection theory revealed trends in the underlying score distribution across subgroups. Specifically, we show variations in the optimal scoring threshold varies for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. 
\end{abstract}

\input{sections/introduction.tex}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
    \caption{\small{\textbf{Quantitative of human assessment.} Different human subgroups listed per row. Each column is the subgroup labeled. Note that each people are best within their subgroup, and second-best within the same subgroup but different gender. CF shows the least variation, but with the lowest accuracy. CM shows the best accuracy, but second to \gls{wm} in deviation from the mean. Thus, scores for males vary more than females.}}
    \label{tab:humsn-eval-results} 
\begin{center}
\footnotesize
\scalebox{0.9}{
\begin{tabular}{l c c c c c}
        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474\pm0.041 \\
        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441\pm0.059 \\
        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485\pm0.062 \\
        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447\pm0.108 \\\midrule
        Avg &  0.433& 0.474&0.477 &0.463 &0.462\pm0.020\\
 \end{tabular}}
 \end{center}
 \vspace{-5mm}
\end{table}

@@

\putbib
\end{bibunit}

% \newpage\newpage
% \col
\begin{bibunit}
\input{supplemental.tex}
\end{bibunit}
% \end{comment}

>>>>

\usepackage{FG2019}
\usepackage{multirow}
\newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}
\usepackage[acronym]{glossaries}
\usepackage{acronym}
\usepackage{bigstrut}
% symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{amsmath}% assumes amsmath package installed
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% \usepackage{hyperref}       % hyperlinks
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

\usepackage{url}            % simple URL typesetting
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
%\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{tikz}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
\usepackage{arydshln}
\usepackage{todonotes}
\usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}

@@

%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage[]{algorithm2e}
%\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont

@@

\newacronym{sdm}{SDM}{signal detection model}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{nmse}{NMSE}{Normalized Mean Square Error}
\newacronym{det}{DET}{Detection Error Trade-off}
\newacronym{tp}{TP}{true-positive}
\newacronym{fp}{FP}{false-positive}

\newacronym{tpir}{TPIR}{true-positive identification rate}
\newacronym{frir}{FRIR}{false-reject identification rate}
\newacronym{fpir}{FRIR}{false-positive identification rate}

@@

\newacronym{tar}{TAR}{True Acceptance Rate}
\newacronym{far}{FAR}{False Acceptance Rate}
\newacronym{eer}{EER}{Equal Error Rate}

@@

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}
\renewcommand{\arraystretch}{1.4}
%\newcommand{\xmark}{\ding{53}}%

@@

@@

%use this in case of several affiliations
\author{\parbox{16cm}{\centering
    {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
    {\normalsize
    \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
    \thanks{}% <-this % stops a space
}

\begin{document}
%\begin{bibunit}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. Classic signal detection theory revealed trends in the underlying score distribution across subgroups. Specifically, we show variations in the optimal scoring threshold varies for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. To download the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}
\
\input{sections/introduction.tex}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
    \caption{\small{\textbf{Quantitative of human assessment.} Different human subgroups listed per row. Each column is the subgroup labeled. Note that each people are best within their subgroup, and second-best within the same subgroup but different gender. CF shows the least variation, but with the lowest accuracy. CM shows the best accuracy, but second to \gls{wm} in deviation from the mean. Thus, scores for males vary more than females.}}
    \label{tab:humsn-eval-results} 
     \vspace{-2mm}

\footnotesize
\scalebox{0.9}{
\begin{tabular}{l c c c c c}
        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474$\pm$0.041 \\
        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441$\pm$0.059 \\
        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485$\pm$0.062 \\
        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447$\pm$0.108 \\\midrule
        Avg &  0.433& 0.474&0.477 &0.463 &0.462$\pm$0.020\\
 \end{tabular}}
 \end{center}
 \vspace{-6mm}
\end{table}

@@

s\bibliographystyle{ieee}
\bibliography{bias}

%\putbib
%\end{bibunit}
%\begin{bibunit}
%\input{supplemental.tex}
%\end{bibunit}

### 2020-02-06 sections/analysis.tex

\section{Results and Analysis}
An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{https://github.com/clcarwin/sphereface_pytorch}{https://github.com/clcarwin/sphereface_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped. The two features were fused by concatenation.

@@

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
 \vspace{-2mm}
\include{tables/ethnicity_far}
\glsreset{tar}
\glsreset{far}
 \vspace{-3mm}

>>>>

\section{Results and Analysis}
An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https:\/\/github.com\/clcarwin\/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped. The two features were fused by concatenation.

@@

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
 \vspace{-2mm}
\begin{center}
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
 \vspace{-3mm}

### 2020-02-06 sections/introduction.tex
updated supplemental citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\glsresetall
\section{Introduction}
As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards are topics of high interest ~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}. One effect of the growing dependency on technology is the increasing concern regarding biased and unfair algorithms.  For instance, \gls{fr} systems can be untrustworthy and racist in some cases~\cite{england2019,snow2018}.

%The trust of people in machinery built to make decisions that range from casual-to-critical matters still takes time to build  .

\begin{figure}[t!]
\centering
         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{images/montage.pdf}  
\caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectfully. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
% \emph{Bottom-left} average per gender.  \emph{Top-left}: average face of \gls{bfw}. Notice the distinguishable variation in face features within and across subgroups. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
% \caption{\small{\textbf{Data of BFW.} Subgroups depicted by average faces. \emph{Top-left}: average of BFW. \emph{Top-right}: averages for races,  Asian (A), Black (B), Indian (I), White (W). \emph{Bottom-left}: averages for genders, Female (F) Male (M). \emph{Bottom-Right}: combined, \gls{af}, \gls{am}, \gls{bf}, \gls{bm}, \gls{if}, \gls{im}, \gls{wf}, \gls{wm}. 
% \emph{Top-left}: average face of \gls{bfw}.
% Notice the distinguishable variation in face features within and between subgroups.}}
\label{fig:avg-faces}
% \vspace{-3pt}
\end{figure}

@@

\begin{table*}[t!]
\scriptsize
    % \glsunset{bfw}
    % \glsunset{rfw}
    % \glsunset{dp}
    \centering
       \caption{\textbf{The proposed \gls{bfw} compared to related datasets.} \gls{bfw} is exactly balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set. Also, \gls{bfw} uses a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation). Furthermore, \gls{rfw} focuses on race-distribution across of pairs, while not considering the distribution of identities. }
    \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
    \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &\approx80,000&\approx12,000& 4 & \xmark & \checkc &\xmark \\
    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
    \end{tabular}

    \label{tab:compared}
    \vspace{-5mm}
\end{table*}

@@

% Fig.~\ref{fig:avg-faces} depicts the average face of each subgroup.

\begin{table*}[!t]
    \centering
    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions. \textit{Top-row:} Statistics of \gls{bfw}. \textit{Middle-row:} Number of pairs for each partition. \textit{Bottom-row:} Accuracy when applying a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup. Columns grouped by race and then further split by gender. Out of millions of pairs, accuracy is inconsistent across subgroups. Furthermore, $F$ tend to perform inferior to that of $M$ (\ie up to 8\%).}}\label{tab:ethnic-splits}
    \include{tables/ethnic_splits}
    %  \vspace{-5mm}t

\end{table*}

The contributions in this paper are as follows:\footnote{Data, code, more visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}.}
\begin{enumerate}
    \item We built a balanced dataset as a proxy to measure verification performance per subgroup.
    \item We analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
    \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    \item We conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
\end{enumerate}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\glsresetall
\section{Introduction}
As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards are topics of high interest ~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}. One effect of the growing dependency on technology is the increasing concern regarding biased and unfair algorithms.  For instance, \gls{fr} systems can be untrustworthy and racist in some cases~\cite{england2019,snow2018}.
.

\begin{figure}[t!]
\centering
         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{images/montage.pdf}  
\caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectfully. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
\label{fig:avg-faces}
% \vspace{-3pt}
\end{figure}

@@

\begin{table*}[t!]
\scriptsize
    \centering
       \caption{\textbf{The proposed \gls{bfw} compared to related datasets.} \gls{bfw} is exactly balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set. Also, \gls{bfw} uses a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation). Furthermore, \gls{rfw} focuses on race-distribution across of pairs, while not considering the distribution of identities. }
    \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
    \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
    \end{tabular}
    \label{tab:compared}
    \vspace{-5mm}
\end{table*}

@@

\begin{table*}[!t]
    \centering
    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions. \textit{Top-row:} Statistics of \gls{bfw}. \textit{Middle-row:} Number of pairs for each partition. \textit{Bottom-row:} Accuracy when applying a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup. Columns grouped by race and then further split by gender. Out of millions of pairs, accuracy is inconsistent across subgroups. Furthermore, $F$ tend to perform inferior to that of $M$ (\ie up to 8\%).}}\label{tab:ethnic-splits}
    \scalebox{1}{
     \resizebox{\textwidth}{!}{%
    \begin{tabular}{r c c c c c c c c l}
        \toprule
        & \multicolumn{2}{c}{Asian (A)} & \multicolumn{2}{c}{Black (B)}  & \multicolumn{2}{c}{Indian (I)}& \multicolumn{2}{c}{White (W)}\\
        \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7}\cmidrule(l){8-9} % spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
         & Female (AF) & Male (AM) & BF & BM& IF & IM & WF & WM&Aggregated\\ % Column names row
        \midrule

       \# Faces  &  2,500&  2,500& 2,500 & 2,500& 2,500 & 2,500 & 2,500 & 2,500 &20,000 \\ % row 1
        \# Subjects & 100& 100& 100  & 100  & 100  & 100& 100 &100&800  \\ % row 2
        \# Faces / Subject  & 25 & 25    & 25 & 25 & 25  & 25  &  25 & 25 & 25\\ % row 3
        % \midrule
\specialrule{.01em}{.05em}{.05em}
            \# Positive Pairs &  30,000&  30,000& 30,000 &30,000 & 30,000 &30,000&30,000 & 30,000 &240,000 \\ % row 1
        \# Negative Pairs & 85,135&  85,232& 85,016  & 85,141  & 85,287  & 85,152& 85,223 &85,193&681,379  \\ % row 2

        \# Pairs (Total) & 115,135 & 115,232    &115016 &115,141 & 115287  & 115,152  &  115,223& 115193 & 921,379\\ % row 3
        % \midrule
        \specialrule{.01em}{.05em}{.05em}
        Acc$@t_g$ & 0.876 & 0.944  &0.934 &0.942 &0.922&0.949 &0.916  &0.918&0.925$\pm$0.022 \\
        $t_o$ & 0.235 &  0.274 & 0.267&0.254  &0.299 & 0.295& 0.242 &0.222&0.261\\ % row 4
        Acc$@t_o$ & 0.916 &0.964 &0.955 &0.971 &0.933 &0.958 & 0.969 &0.973 & 0.955 $\pm$ 0.018\\
        \bottomrule
    \end{tabular}}
    \glsunset{af}
    }
    \glsunset{af}
    \glsunset{am}
    \glsunset{bf}
    \glsunset{bm}
    \glsunset{if}
    \glsunset{im}
    \glsunset{wf}
    \glsunset{wm}
    \vspace{-12pt}
\end{table*}

The contributions in this paper are as follows:
\begin{enumerate}
    \item We built a balanced dataset as a proxy to measure verification performance per subgroup.
    \item We analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
    \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    \item We conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
\end{enumerate}

### 2020-02-06 sections/relatedworks.tex
updated supplemental citations

% \vspace{-5pt}
\subsection{Bias in \gls{fr}}
Problems of bias in \gls{fr} have been driven by different motivations and have thereforem been solved by different means. To name a few: problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even exploratory data analysis~\cite{muthukumar2019}. Yin et al. \cite{yin2019feature} proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others. Similarly,~\cite{ding2018one} formulated the imbalanced class as one-shot learning and trained a \gls{gan} to generate face features as a means of augmenting class with as few as one sample. \cite{huang2018generative} proposed \gls{gapf} to create fair representations of the data in a quantifiable way, allowing for the finding of a decorrelation scheme from the data without access to its statistics. Wang et al. \cite{wang2018they} defined subgroups at a finer-level (\ie Chinese, Japanese, Korean) to determine the familiarity of faces across these subgroups. Genders have also been used as subgroups for work in bias, whether for efforts of analysis and understanding of gender-based face encodings~\cite{muthukumar2019}. Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the gap between races by knowledge transfer, which was supported by a novel labeled data collection, \gls{rfw}. The release of \gls{rfw} came after the completion of \gls{bfw}. Although similar in terms of demographics (\ie Asian, Black, Indian, and White), \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} based on the assumption CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} was used for training. In contrast, our \gls{bfw} samples from VGG2 and assumes MSCeleb was the training set. As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. For this, we expect bias from data, for this is the training set. Furthermore, our experiments allow for the parsing of results based on gender demographics (\ie although \gls{rfw} does separate males and females in each subgroup, no consideration for splits in just gender or just races are incorporated into the experimental design). Thus, \gls{rfw} and \gls{bfw} are complementary, with both adding metadata on demographics for subjects with faces in renowned, large-scale \gls{fr} datasets.

@@

% \vspace{-5pt}
\subsection{Human bias in \gls{ml}}
Bias is not unique to \gls{ml}, as humans also susceptible to a perceived bias across demographics: it exists across different races, genders, and even ages~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}. \cite{wang2018they} showed machines surpass human performance in classifying faces as Japanese, Chinese, or Korean by nearly 150\%. Precisely, humans barely pass random with 
38.89\% accuracy (\ie $\frac{1}{3}$ is random).

>>>>

\subsection{Bias in \gls{fr}}
Problems of bias in \gls{fr} have been driven by different motivations and have therefore been solved by different means. To name a few: problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even exploratory data analysis~\cite{muthukumar2019}. Yin et al. \cite{yin2019feature} proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others. Similarly,~\cite{ding2018one} formulated the imbalanced class as one-shot learning and trained a \gls{gan} to generate face features as a means of augmenting class with as few as one sample. \cite{huang2018generative} proposed \gls{gapf} to create fair representations of the data in a quantifiable way, allowing for the finding of a decorrelation scheme from the data without access to its statistics. Wang et al. \cite{wang2018they} defined subgroups at a finer-level (\ie Chinese, Japanese, Korean) to determine the familiarity of faces across these subgroups. Genders have also been used as subgroups for work in bias, whether for efforts of analysis and understanding of gender-based face encodings~\cite{muthukumar2019}. Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the gap between races by knowledge transfer, which was supported by a novel labeled data collection, \gls{rfw}. The release of \gls{rfw} came after the completion of \gls{bfw}. Although similar in terms of demographics (\ie Asian, Black, Indian, and White), \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} based on the assumption CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} was used for training. In contrast, our \gls{bfw} samples from VGG2 and assumes MSCeleb was the training set. As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. For this, we expect bias from data, for this is the training set. Furthermore, our experiments allow for the parsing of results based on gender demographics (\ie although \gls{rfw} does separate males and females in each subgroup, no consideration for splits in just gender or just races are incorporated into the experimental design). Thus, \gls{rfw} and \gls{bfw} are complementary, with both adding metadata on demographics for subjects with faces in renowned, large-scale \gls{fr} datasets.

@@

\subsection{Human bias in \gls{ml}}
Bias is not unique to \gls{ml}, as humans also susceptible to a perceived bias across demographics: it exists across different races, genders, and even ages~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}. \cite{wang2018they} showed machines surpass human performance in classifying faces as Japanese, Chinese, or Korean by nearly 150\%. Precisely, humans barely pass random with 
38.89\% accuracy (\ie $\frac{1}{3}$ is random).
